% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{svg}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,
    urlcolor=blue,
}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.


% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\begin{document}
%
\title{Detecting life-threatening patterns in Point-of-care ECG using efficient memory and processor power\thanks{Supervisor: Professor Pedro Pereira Rodrigues (MEDCIDS, Porto), co-supervisor: Professor Eamonn Keogh (UCR, Riverside)}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
\titlerunning{ ECG using efficient memory and processor power }
%
%
\author{  Francisco Bischoff\inst{1}\orcidID{0000-0002-5301-8672} }

\authorrunning{ Bischoff F. }
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{  \textsuperscript{1} Department of Community Medicine, Information and Health Decision Sciences (MEDCIDS), Faculty of
Medicine, University of Porto, Porto, Portugal }

% \institute{true}

\maketitle              % typeset the header of the contribution
%

\begin{abstract}
  Currently, Point-of-Care (POC) ECG monitoring works either as plot devices or alarms for abnormal
  cardiac rhythms using predefined normal trigger ranges. Great effort has been made to improve the
  accuracy of such monitoring, but in the ICU setting. Thinking outside the ICU setting, where
  high-end devices are available, we aim to identify, on streaming data, life-threatening hearth
  electric patterns using low CPU and memory, enabling ward monitors, home devices and even wearable
  devices to be able to identify such events.

  \keywords{
        anomaly detection \and
        ECG \and
        matrix profile \and
        time series \and
        point-of-care  }

\end{abstract}
%
%
% \def\spacingset#1{\renewcommand{\baselinestretch}%
% {#1}\small\normalsize} \spacingset{1}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Currently, Point-of-Care (POC) ECG monitoring works either as plot devices or alarms for abnormal
cardiac rhythms using predefined normal trigger ranges. Modern devices also incorporate algorithms
to analyze arrhythmias improving their specificity. On the other hand, full 12-derivation ECG
machines are complex, are not suited to use as simple monitors, and are used with strict techniques
for formal diagnostics of hearth electric conduction pathologies. The automatic diagnostics are
derived from a complete analysis of the 12-dimension data after it is fully and well collected.

In February of 2015, the CinC/Physionet Challenge 2015 was about ``Reducing False Arrhythmia Alarms
in the ICU'' {[}\protect\hyperlink{ref-Clifford2015}{2}{]}. The introduction article stated that it had been reported that up to 86\%
resulting of the alarms are false, and this can lead to decreased staff attention and an increase in
patients' delirium {[}\protect\hyperlink{ref-Chambrin2001}{1}, \protect\hyperlink{ref-Lawless1994}{10}, \protect\hyperlink{ref-Parthasarathy2004}{13}{]}.

This subject draws attention to the importance of correctly identify abnormal hearth electric
patterns. Meanwhile, this opens the opportunity of thinking outside the ICU setting, where we still
monitoring patients (and ourselves) using devices with low processing power, as for example ward
monitors, home devices and wearable devices.

\hypertarget{objectives-and-research-question}{%
\section{Objectives and research question}\label{objectives-and-research-question}}

While this research was inspired on the Physionet's challenge, its purpose is not to beat the state
of the art on that challenge, but to identify, on streaming data, abnormal hearth electric patterns,
specifically those which are life-threatening, using low CPU and low memory requirements.

The main questions is: can we accomplish this objective using a minimalist approach (low CPU, low
memory) while maintaining robustness?

\hypertarget{related-works}{%
\section{Related Works}\label{related-works}}

The Physionet's challenge yielded several papers on the subject
{[}\protect\hyperlink{ref-couto2015}{3}, \protect\hyperlink{ref-fallet2015}{5}, \protect\hyperlink{ref-hoogantink2015}{8}, \protect\hyperlink{ref-kalidas2015}{9}, \protect\hyperlink{ref-plesinger2015}{14}{]}. However, independently of
their approach to this problem, none of the authors reported benchmarks, memory usage, robustness
test, or context invariance that could assure its implementation smaller devices.

\hypertarget{planned-approach}{%
\section{Planned approach}\label{planned-approach}}

\hypertarget{the-data}{%
\subsection{The data}\label{the-data}}

The dataset used is the CinC/Physionet Challenge 2015 public dataset {[}\protect\hyperlink{ref-Clifford2015}{2}{]}, composed of
750 patients with at least five minutes records. The \emph{events} we seek to identify are the
life-threatening arrhythmias as defined by Physionet's challenge {[}\protect\hyperlink{ref-Clifford2015}{2}{]}.

\hypertarget{matrix-profile}{%
\subsection{Matrix Profile}\label{matrix-profile}}

This work will use the state-of-the-art {[}\protect\hyperlink{ref-DePaepe2020}{4}, \protect\hyperlink{ref-Feremans2020}{6}{]} time series analysis technique
called Matrix Profile (MP) that once computed, allows us to derive frameworks to all sorts of tasks,
as motif discovery, anomaly detection, regime change detection and others {[}\protect\hyperlink{ref-Yeh2017a}{16}{]}. The MP is
known to be an incredible fast algorithm {[}\protect\hyperlink{ref-gharghabi2018}{7}, \protect\hyperlink{ref-Madrid2019}{11}{]}, thus viewing on the other
side (not to process billions of data points using a desktop), it has a great potential to be used
on small devices.

The streaming data, coming from one patient, is processed to create its MP in real-time. Then, the
FLOSS algorithm {[}\protect\hyperlink{ref-gharghabi2018}{7}{]} is computed for detecting a regime change. When a new regime is
detected, a sample of this new regime is analysed by a model and a decision is made. If the new
regime is life-threatening, the alarm will be fired.

\hypertarget{detecting-regime-changes}{%
\subsection{Detecting regime changes}\label{detecting-regime-changes}}

The regime change detection will be using the FLOSS algorithm {[}\protect\hyperlink{ref-gharghabi2018}{7}{]} which is an on-line
algorithm built on top of the computed MP. The algorithm is based on the assumption that between
two regimes, the most similar shape (its nearest neighbor, 1-NN) is located on ``the same side''. The
details of the algorithm are described in the original paper {[}\protect\hyperlink{ref-gharghabi2018}{7}{]}. In short, the
algorithm keeps track of the number of 1-NN references (called \emph{Arc Counts}) that crosses a point in
time. As shown on the original article, figures 2 and 3 {[}\protect\hyperlink{ref-gharghabi2018}{7}{]}, a drop in the Arc Counts
indicates a regime change.

\hypertarget{classification-of-the-new-regime}{%
\subsection{Classification of the new regime}\label{classification-of-the-new-regime}}

The next step is to verify if the new regime detected is indeed a life-threatening pattern. The
method of choice is a classification model using shapelets as signatures of such patterns. The aim is
not to identify the exact type of the new regime, but if it is life-threatening or not. This allows
us to use a set of shapelet candidates which maximize the classification performance.

Leveraging on the MP concept, we can use the Contrast Profile (CP) {[}\protect\hyperlink{ref-Mercer2021}{12}{]} to derive a set of
shapelets candidates. The CP looks for patterns that are at the same time very \emph{similar} to its
neighbors in class \emph{A} while is very \emph{different} from the nearest neighbor from class \emph{B}.

For a more complete understanding of the process, in the original article, the figure 6 shows a
practical example {[}\protect\hyperlink{ref-Mercer2021}{12}{]}.

\hypertarget{implementation}{%
\subsection{Implementation}\label{implementation}}

Ultimately, this workflow will be experimented on a low power device, such as an ESP32 microcontroller {[}\protect\hyperlink{ref-esp32}{15}{]}
in order to validate the concept and measure the performance and benchmarks.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Chambrin2001}{}}%
\CSLLeftMargin{1. }
\CSLRightInline{Chambrin, M.C.: Alarms in the intensive care unit: How can the number of false alarms be reduced? Critical care (London, England). 5, 4, 184--8 (2001). https://doi.org/\href{https://doi.org/10.1186/cc1021}{10.1186/cc1021}.}

\leavevmode\vadjust pre{\hypertarget{ref-Clifford2015}{}}%
\CSLLeftMargin{2. }
\CSLRightInline{Clifford, G.D. et al.: The PhysioNet/computing in cardiology challenge 2015: Reducing false arrhythmia alarms in the ICU. In: Computing in cardiology. (2015). https://doi.org/\href{https://doi.org/10.1109/cic.2015.7408639}{10.1109/cic.2015.7408639}.}

\leavevmode\vadjust pre{\hypertarget{ref-couto2015}{}}%
\CSLLeftMargin{3. }
\CSLRightInline{Couto, P. et al.: Suppression of false arrhythmia alarms using ECG and pulsatile waveforms. Presented at the September (2015). https://doi.org/\href{https://doi.org/10.1109/cic.2015.7411019}{10.1109/cic.2015.7411019}.}

\leavevmode\vadjust pre{\hypertarget{ref-DePaepe2020}{}}%
\CSLLeftMargin{4. }
\CSLRightInline{De Paepe, D. et al.: {A generalized matrix profile framework with support for contextual series analysis}. Engineering Applications of Artificial Intelligence. 90, January, 103487 (2020). https://doi.org/\href{https://doi.org/10.1016/j.engappai.2020.103487}{10.1016/j.engappai.2020.103487}.}

\leavevmode\vadjust pre{\hypertarget{ref-fallet2015}{}}%
\CSLLeftMargin{5. }
\CSLRightInline{Fallet, S. et al.: A multimodal approach to reduce false arrhythmia alarms in the intensive care unit. Presented at the September (2015). https://doi.org/\href{https://doi.org/10.1109/cic.2015.7408640}{10.1109/cic.2015.7408640}.}

\leavevmode\vadjust pre{\hypertarget{ref-Feremans2020}{}}%
\CSLLeftMargin{6. }
\CSLRightInline{Feremans, L. et al.: {Pattern-Based Anomaly Detection in Mixed-Type Time Series}. In: Lecture notes in computer science (including subseries lecture notes in artificial intelligence and lecture notes in bioinformatics). pp. 240--256 (2020). https://doi.org/\href{https://doi.org/10.1007/978-3-030-46150-8_15}{10.1007/978-3-030-46150-8\_15}.}

\leavevmode\vadjust pre{\hypertarget{ref-gharghabi2018}{}}%
\CSLLeftMargin{7. }
\CSLRightInline{Gharghabi, S. et al.: Domain agnostic online semantic segmentation for multi-dimensional time series. Data Mining and Knowledge Discovery. 33, 1, 96--130 (2018). https://doi.org/\href{https://doi.org/10.1007/s10618-018-0589-3}{10.1007/s10618-018-0589-3}.}

\leavevmode\vadjust pre{\hypertarget{ref-hoogantink2015}{}}%
\CSLLeftMargin{8. }
\CSLRightInline{Hoog Antink, C., Leonhardt, S.: Reducing false arrhythmia alarms using robust interval estimation and machine learning. Presented at the September (2015). https://doi.org/\href{https://doi.org/10.1109/cic.2015.7408642}{10.1109/cic.2015.7408642}.}

\leavevmode\vadjust pre{\hypertarget{ref-kalidas2015}{}}%
\CSLLeftMargin{9. }
\CSLRightInline{Kalidas, V., Tamil, L.S.: Enhancing accuracy of arrhythmia classification by combining logical and machine learning techniques. Presented at the September (2015). https://doi.org/\href{https://doi.org/10.1109/cic.2015.7411015}{10.1109/cic.2015.7411015}.}

\leavevmode\vadjust pre{\hypertarget{ref-Lawless1994}{}}%
\CSLLeftMargin{10. }
\CSLRightInline{Lawless, S.T.: Crying wolf: False alarms in a pediatric intensive care unit. Critical care medicine. 22, 6, 981--5 (1994). https://www.ncbi.nlm.nih.gov/pubmed/\href{https://www.ncbi.nlm.nih.gov/pubmed/8205831}{8205831}.}

\leavevmode\vadjust pre{\hypertarget{ref-Madrid2019}{}}%
\CSLLeftMargin{11. }
\CSLRightInline{Madrid, F. et al.: {Matrix Profile XVI: Efficient and Effective Labeling of Massive Time Series Archives}. In: 2019 IEEE international conference on data science and advanced analytics (DSAA). pp. 463--472 IEEE (2019). https://doi.org/\href{https://doi.org/10.1109/DSAA.2019.00061}{10.1109/DSAA.2019.00061}.}

\leavevmode\vadjust pre{\hypertarget{ref-Mercer2021}{}}%
\CSLLeftMargin{12. }
\CSLRightInline{Mercer, R. et al.: Matrix profile XXIII: Contrast profile: A novel time series primitive that allows real world classification. In: 2021 IEEE international conference on data mining (ICDM). pp. 1240--1245 (2021). https://doi.org/\href{https://doi.org/10.1109/ICDM51629.2021.00151}{10.1109/ICDM51629.2021.00151}.}

\leavevmode\vadjust pre{\hypertarget{ref-Parthasarathy2004}{}}%
\CSLLeftMargin{13. }
\CSLRightInline{Parthasarathy, S., Tobin, M.J.: Sleep in the intensive care unit. Intensive Care Medicine. 30, 2, 197--206 (2004). https://doi.org/\href{https://doi.org/10.1007/s00134-003-2030-6}{10.1007/s00134-003-2030-6}.}

\leavevmode\vadjust pre{\hypertarget{ref-plesinger2015}{}}%
\CSLLeftMargin{14. }
\CSLRightInline{Plesinger, F. et al.: False alarms in intensive care unit monitors: Detection of life-threatening arrhythmias using elementary algebra, descriptive statistics and fuzzy logic. Presented at the September (2015). https://doi.org/\href{https://doi.org/10.1109/cic.2015.7408641}{10.1109/cic.2015.7408641}.}

\leavevmode\vadjust pre{\hypertarget{ref-esp32}{}}%
\CSLLeftMargin{15. }
\CSLRightInline{Wikipedia: ESP32, \url{https://en.wikipedia.org/wiki/ESP32}, last accessed 2022/06/13.}

\leavevmode\vadjust pre{\hypertarget{ref-Yeh2017a}{}}%
\CSLLeftMargin{16. }
\CSLRightInline{Yeh, C.-C.M. et al.: Matrix profile i: All pairs similarity joins for time series: A unifying view that includes motifs, discords and shapelets. In: 2016 IEEE 16th international conference on data mining (ICDM). pp. 1317--1322 IEEE (2016). https://doi.org/\href{https://doi.org/10.1109/ICDM.2016.0179}{10.1109/ICDM.2016.0179}.}

\end{CSLReferences}

\end{document}
